---
title: "612 project"
output:
  html_document:
    toc: false
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    toc: false
    number_sections: false
---

# 612 Time Series Project

Package requirements:

```{r setup, message=FALSE, warning=FALSE}
# install.packages("quantmod")
# install.packages("fredr")
# install.packages("ggfortify")
# install.packages('urca')
# install.packages("tseries")
# install.packages("forecast")
# install.packages("dynlm")
# install.packages("stargazer")
# install.packages("pracma")
# install.packages("dLagM")
# install.packages("gets")
# install.packages("car")
# install.packages("lmtest")
# install.packages("vars")
# install.packages("devtools", type = "binary")
# install.packages("tsDyn")
# install.packages('fGarch', type = "binary")
# No extra package needed: STL adjustment uses base R stats::stl()

options(warn=-1)
```

Library requirements:

```{r}
library(fGarch)
library(xts)
library(urca)
library(stargazer)
library(ggplot2)
library(ggfortify)
library(scales)
library(patchwork)
library(forecast)
library(aTSA)
library(rugarch)
library(lubridate)
library(vars)
# seasonal adjustment uses base R stats::stl() — no extra package needed
```

------------------------------------------------------------------------

## Part 1: Data Import and Cleaning

All datasets are imported and preprocessed here before any analysis begins. This includes frequency alignment, missing value handling, and seasonal adjustment of the trade series.

### 1.1 EUR/CNY Exchange Rate (Daily)

```{r}
# Import EUR/CNY exchange rate (daily frequency)
fx = read.csv('https://raw.githubusercontent.com/Yumeng20021214/612-project/refs/heads/Yumeng/exchange_rate_final.csv')

# Convert TIME_PERIOD to Date format
fx$TIME_PERIOD = as.Date(fx$TIME_PERIOD, format = "%m/%d/%Y")
xts_fx = xts(fx$OBS_VALUE, order.by = fx$TIME_PERIOD)
colnames(xts_fx) = "fx"

cat("FX sample: ", format(index(xts_fx)[1]), "to",
    format(tail(index(xts_fx), 1)), "\n")
cat("Observations:", nrow(xts_fx), "\n")
tail(xts_fx)
```

### 1.2 China–EU Trade (Monthly)

```{r}
# Import monthly export data
trade = read.csv("https://raw.githubusercontent.com/Yumeng20021214/612-project/refs/heads/Yumeng/export_final.csv")

# Remove missing values (marked as ":")
trade = trade[trade$Export != ":", ]

# Convert date and numeric; align index to end-of-month
trade$Month  = as.Date(paste0(trade$Month, "-01"))
trade$Month  = ceiling_date(trade$Month, "month") - days(1)
trade$Export = as.numeric(trade$Export)
xts_trade    = xts(trade$Export, order.by = trade$Month)
colnames(xts_trade) = "trade"

cat("Trade sample: ", format(index(xts_trade)[1]), "to",
    format(tail(index(xts_trade), 1)), "\n")
cat("Observations:", nrow(xts_trade), "\n")
tail(xts_trade)
```

### 1.3 VIX — CBOE Volatility Index (Daily → Monthly)

```{r}
# Import VIX (CBOE Volatility Index, daily) from FRED
# Series: VIXCLS — available from 1990-01-02, daily
vix_raw        = read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=VIXCLS",
                           stringsAsFactors = FALSE)
colnames(vix_raw) = c("DATE", "VIXCLS")
vix_raw$DATE   = as.Date(vix_raw$DATE)
vix_raw$VIXCLS = suppressWarnings(as.numeric(vix_raw$VIXCLS))
vix_raw        = vix_raw[!is.na(vix_raw$DATE) & !is.na(vix_raw$VIXCLS), ]

# Average to monthly and align index to end-of-month
xts_vix_daily   = xts(vix_raw$VIXCLS, order.by = vix_raw$DATE)
xts_vix_monthly = apply.monthly(xts_vix_daily, colMeans)
index(xts_vix_monthly) = ceiling_date(index(xts_vix_monthly), "month") - days(1)
colnames(xts_vix_monthly) = "vix"

cat("VIX sample: ", format(index(xts_vix_monthly)[1]), "to",
    format(tail(index(xts_vix_monthly), 1)), "\n")
cat("Observations:", nrow(xts_vix_monthly), "\n")
tail(xts_vix_monthly)
```

### 1.4 Euro Area Industrial Production (Monthly)

```{r}
# Import Euro area industrial production index (monthly) from FRED
# Series: EA19PRINTO01IXOBSAM — index 2015=100, seasonally adjusted by OECD
eaip_raw        = read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=EA19PRINTO01IXOBSAM",
                            stringsAsFactors = FALSE)
colnames(eaip_raw) = c("DATE", "EAIP")
eaip_raw$DATE   = as.Date(eaip_raw$DATE)
eaip_raw$EAIP   = suppressWarnings(as.numeric(eaip_raw$EAIP))
eaip_raw        = eaip_raw[!is.na(eaip_raw$DATE) & !is.na(eaip_raw$EAIP), ]

# Align to end-of-month (FRED reports monthly obs on the 1st)
eaip_raw$DATE   = ceiling_date(eaip_raw$DATE, "month") - days(1)
xts_eaip        = xts(eaip_raw$EAIP, order.by = eaip_raw$DATE)
colnames(xts_eaip) = "eaip"

cat("Euro area IP sample: ", format(index(xts_eaip)[1]), "to",
    format(tail(index(xts_eaip), 1)), "\n")
cat("Observations:", nrow(xts_eaip), "\n")
tail(xts_eaip)
```

### 1.5 Aggregate to Monthly and Merge All Series

```{r}
# Convert daily exchange rate level to monthly average; align to end-of-month
level_monthly = apply.monthly(xts_fx, colMeans)
index(level_monthly) = ceiling_date(index(level_monthly), "month") - days(1)

# Merge exchange rate level, raw trade, VIX, and Euro area IP
data_raw = merge(level_monthly, xts_trade, xts_vix_monthly, xts_eaip)
data_raw = na.omit(data_raw)
colnames(data_raw) = c("level", "trade", "vix", "eaip")

head(data_raw)
tail(data_raw)
cat("Merged sample:", format(index(data_raw)[1]), "to",
    format(tail(index(data_raw), 1)), "\n")
cat("Observations:", nrow(data_raw), "\n")
```

### 1.6 Seasonal Adjustment of Trade Data

Chinese export data exhibits very strong calendar effects (e.g. Chinese New Year in January/February causes swings of up to ±11 percentage points in the raw month-on-month growth rate). Failing to remove this seasonal pattern leaves large predictable fluctuations in the residuals, which invalidates Portmanteau tests, Granger causality, and IRF inference.

We use **STL decomposition** (Seasonal-Trend decomposition using Loess, Cleveland et al. 1990) implemented in base R via `stl()`. STL is a robust, widely-used method that handles Chinese calendar effects well with a flexible seasonal window.

```{r}
# Convert trade to a monthly ts object for seasonal adjustment
trade_ts = ts(
  as.numeric(data_raw$trade),
  start     = c(year(index(data_raw)[1]), month(index(data_raw)[1])),
  frequency = 12
)

# STL decomposition (s.window = "periodic" assumes a stable seasonal pattern)
stl_fit = stl(log(trade_ts), s.window = "periodic", robust = TRUE)

# Seasonally adjusted series = original - seasonal component (in log space)
log_trade_sa = log(trade_ts) - stl_fit$time.series[, "seasonal"]
trade_sa_ts  = exp(log_trade_sa)

# Convert back to xts (same index as data_raw)
trade_sa_xts = xts(as.numeric(trade_sa_ts), order.by = index(data_raw))
colnames(trade_sa_xts) = "trade_sa"

# Visually inspect original vs. adjusted
trade_compare_df = data.frame(
  Date     = index(data_raw),
  Original = as.numeric(data_raw$trade),
  Adjusted = as.numeric(trade_sa_xts)
)

ggplot(trade_compare_df, aes(x = Date)) +
  geom_line(aes(y = log(Original), color = "Original (log)"),  linewidth = 0.6) +
  geom_line(aes(y = log(Adjusted), color = "SA (log)"),        linewidth = 0.6) +
  scale_color_manual(values = c("Original (log)" = "grey60", "SA (log)" = "seagreen4")) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    legend.title    = element_blank(),
    legend.position = "top",
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Log Monthly Export") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# Verify: seasonal pattern is removed in growth rates
dlogtrade_raw_check = diff(log(trade_ts)) * 100
dlogtrade_sa_check  = diff(log(trade_sa_ts)) * 100

cat("=== Raw dlogtrade: mean by calendar month ===\n")
print(round(tapply(as.numeric(dlogtrade_raw_check),
                   cycle(dlogtrade_raw_check), mean), 2))

cat("\n=== SA dlogtrade: mean by calendar month ===\n")
print(round(tapply(as.numeric(dlogtrade_sa_check),
                   cycle(dlogtrade_sa_check), mean), 2))
```

#### Interpretation

The seasonal means table confirms that before adjustment the monthly growth rates vary from about −11% (February) to +11% (January), a range of 22 percentage points driven entirely by the Chinese New Year. After STL adjustment, the means by calendar month are all close to zero, confirming that the seasonal pattern has been successfully removed. All subsequent analysis uses the SA trade series.

```{r}
# Replace raw trade with SA trade in the merged dataset
data_all = merge(data_raw$level, trade_sa_xts, data_raw$vix, data_raw$eaip)
data_all = na.omit(data_all)
colnames(data_all) = c("level", "trade_sa", "vix", "eaip")

cat("Final merged sample:", format(index(data_all)[1]), "to",
    format(tail(index(data_all), 1)), "\n")
cat("Observations:", nrow(data_all), "\n")
```

------------------------------------------------------------------------

## Part 2: EUR/CNY Exchange Rate — Individual Analysis

### 2.1 Daily Returns: Computation and Stationarity

```{r}
# Compute log returns (first difference of log exchange rate)
dlogfx = diff(log(xts_fx))
rfx = na.omit(dlogfx)
head(rfx)
tail(rfx)
```

```{r}
# ADF test on daily returns
summary(ur.df(rfx, type = 'trend', lags = 8, selectlags = "AIC"))
```

#### Interpretation of the unit root test

Test statistics ( type = "trend" )

$\phi_3$ = 1379.367 vs. 5% critical value = 6.25 $\to$ Reject $H_0$ (deterministic trend = 0) $\to$ a deterministic trend should be included.

$\tau_3$ = -52.5236 vs. 5% critical value = -3.41 $\to$ Reject $H_0$ (no unit root).

Conclusion: The specification including a deterministic trend is appropriate and provides evidence of no unit root in EUR/CNY exchange rate returns.

```{r}
# Regress return on a deterministic time trend
t = 1:length(rfx)
model = lm(rfx ~ t)
stargazer(model, type = "text")
```

#### Interpretation of the trend regression

Model: $y_t$ = $\alpha$ + $\beta_t$ + $\varepsilon_t$

The estimated trend coefficient ($\beta$) is approximately 0.00000. The coefficient is not statistically significant (p $>$ 0.1). This means there is no significant deterministic trend in the return series.

Conclusion: The EUR/CNY exchange rate returns are stationary with no deterministic trend to remove. The mean of returns remains constant over time, and any fluctuations are random rather than systematic.

### 2.2 Time Series Plot of Daily Returns

```{r}
rfx_df = data.frame(
  Date   = index(rfx),
  Return = as.numeric(rfx)
)

ggplot(rfx_df, aes(x = Date, y = Return)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "EUR/CNY Exchange Rate Return") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)
```

#### Interpretation of volatility patterns

The figure displays daily EUR/CNY exchange rate returns over a long sample period. The series fluctuates around zero, but the size of the fluctuations changes substantially over time, indicating that volatility is not constant.

2005-2006: Volatility in 2005–2006 was relatively moderate, but it was influenced by important structural changes in China's exchange-rate regime. In 2005, China moved away from the long-standing strict U.S. dollar peg and began shifting toward a more market-based, managed floating system, which increased the role of market forces in RMB pricing.

2008–2009: Volatility rises sharply during the global financial crisis, with much larger positive and negative movements than in earlier years. This is one of the most turbulent periods, and several extreme spikes are visible.

2010–2012: Volatility declines compared with the crisis period, but fluctuations remain more pronounced than in the pre-2008 years, indicating that market uncertainty is still present.

2015–2016: Volatility rose markedly, largely because of China's exchange-rate reform, concerns about China's slowing growth, capital outflow pressures, and uncertainty in global financial markets.

Around 2022: Volatility increased again, mainly because of the conflict between Russia and Ukraine, the resulting energy shock in Europe, and heightened inflation and policy uncertainty.

Around 2025: Volatility after 2025 can be linked to elevated trade-policy uncertainty, tariff shocks, and broader geopolitical tensions.

Overall, the plot suggests that EUR/CNY exchange rate volatility is not constant, but instead tends to increase during periods of economic or financial stress. The visible clustering of large and small movements provides preliminary evidence in favor of using a GARCH-type model to capture time-varying volatility.

### 2.3 Testing for ARCH Effects

#### Method A: Plot of the squared series

```{r}
rfx2_df = data.frame(
  Date            = index(rfx),
  Squared_Returns = as.numeric(rfx^2)
)

ggplot(rfx2_df, aes(x = Date, y = Squared_Returns)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Squared Returns") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)
```

#### Interpretation of the plot of squared returns

We observe moderate EUR/CNY return volatility during the early part of the sample. However, volatility increases noticeably during several periods, especially around 2008–2009, 2015–2016, and 2025, where large spikes appear more frequently, indicating episodes of heightened exchange-rate uncertainty.

Additional periods of elevated volatility are visible in 2010–2012. Other noticeable increases appear around 2020, likely reflecting the disruption caused by the COVID-19 pandemic, and around 2022, indicating the impact of the Russia-Ukraine conflict. Outside these episodes, volatility tends to be lower, with squared returns clustering closer to zero.

Overall, the plot suggests the presence of volatility clustering, providing preliminary evidence in favor of using a GARCH-type model to capture time-varying volatility.

#### Method B: ACF plot of the squared series

```{r}
options(repr.plot.width=8, repr.plot.height=8)
acf(rfx^2)
```

#### Interpretation of the ACF plot of squared returns

The ACF of squared returns shows positive autocorrelations at the first several lags, and many of these values remain above the confidence bands. This pattern indicates that dependence in the squared returns is persistent rather than short-lived, providing evidence of volatility clustering consistent with ARCH/GARCH-type effects.

#### Method C: Plot of the squared residuals

#### Estimate conditional mean model

```{r}
auto.arima(rfx,
           max.d      = 0,
           max.p      = 20,
           max.q      = 20,
           seasonal   = FALSE,
           ic         = "aic",
           stationary = TRUE,
           stepwise   = FALSE
)
```

```{r}
# Estimate AR(1) and extract residuals
model.rfx = arima(rfx, order = c(1, 0, 0))
res.rfx   = model.rfx$residuals
res.rfx   = xts(res.rfx, order.by = index(rfx))
```

```{r}
# Squared residuals
rfx3_df = data.frame(
  Date              = index(rfx),
  Squared_Residuals = as.numeric(res.rfx^2)
)

ggplot(rfx3_df, aes(x = Date, y = Squared_Residuals)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Squared Residuals") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)
```

#### Interpretation of the plot of squared residuals

The squared residuals closely resemble the squared returns, indicating that the estimated mean model (AR(1)) removes very little variation from the original series. This suggests that the predictable component in the mean is small, and most of the dynamics are driven by the innovation term.

#### Method D: ACF of the squared residuals

```{r}
options(repr.plot.width=8, repr.plot.height=8)
acf(res.rfx^2)
```

#### Interpretation of the ACF plot of squared residuals

The ACF of squared residuals shows positive autocorrelations at the first several lags that remain above the confidence bands. This pattern is very similar to the ACF of squared returns, suggesting that modeling the mean equation does not remove the volatility dependence. This provides support for ARCH/GARCH-type effects.

#### Method E: Engle's ARCH-LM Test

```{r}
aTSA::arch.test(model.rfx)
```

#### Interpretation of ARCH test results

All p-values are equal to 0 for both the Portmanteau-Q and LM tests. The null hypothesis of no ARCH effects is strongly rejected at every lag. A GARCH-type model is clearly warranted.

### 2.4 Estimating ARMA-GARCH Model

```{r}
spec = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "norm"
)
garch.dfx = ugarchfit(spec, rfx)
garch.dfx@fit$matcoef
infocriteria(garch.dfx)
```

#### Interpretation of the AR(1)-GARCH(1,1) Model

The AR(1) coefficient ($ar1 = -0.037^{***}$) is negative and statistically significant, indicating a small degree of short-run mean reversion. The ARCH ($\alpha_1 = 0.032^{***}$) and GARCH ($\beta_1 = 0.965^{***}$) coefficients are both highly significant, confirming strong and persistent volatility clustering.

#### Removing mean and re-estimating ARMA-GARCH model

```{r}
spec2 = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = FALSE),
  distribution.model = "norm"
)
garch.dfx2 = ugarchfit(spec2, rfx)
garch.dfx2@fit$matcoef
infocriteria(garch.dfx2)
```

Results are nearly unchanged. For simplicity, we proceed with the zero-mean specification.

#### Using Student's t-distribution in place of the normal distribution

```{r}
spec3 = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = FALSE),
  distribution.model = "std"
)
garch.dfx3 = ugarchfit(spec3, rfx)
garch.dfx3@fit$matcoef
infocriteria(garch.dfx3)
```

#### Interpretation of the AR(1)-GARCH(1,1) Model with Student's t Distribution

The shape parameter (6.19\*\*\*) is statistically significant, confirming heavy tails. The lower information criteria relative to the normal-error specification confirm that the Student's t distribution provides a better fit.

#### Residual Diagnostics

```{r}
Box.test(residuals(garch.dfx3, standardize = TRUE),   lag = 8, type = "Ljung-Box")
Box.test(residuals(garch.dfx3, standardize = TRUE)^2, lag = 8, type = "Ljung-Box")
```

Both p-values are well above 0.05, indicating no remaining serial dependence in residuals or squared residuals. The AR(1)-GARCH(1,1) model with Student's t errors is adequate.

### 2.5 Extracting and Re-scaling Conditional Volatility

The GARCH conditional variance is on the order of $10^{-6}$, while exchange-rate growth and trade growth are measured in percentage points (order 1–10). Including variables with such extreme scale differences in a VAR distorts the coefficient estimates and makes impulse responses economically uninterpretable.

We convert the conditional variance to **conditional standard deviation** and express it in percentage points (multiply by 100), so all VAR variables share the same order of magnitude.

```{r}
# Conditional standard deviation in percentage points (daily)
vol_daily_sd = sigma(garch.dfx3) * 100

# Convert to monthly average; align index to end-of-month
vol_monthly_sd = apply.monthly(vol_daily_sd, colMeans)
index(vol_monthly_sd) = ceiling_date(index(vol_monthly_sd), "month") - days(1)
colnames(vol_monthly_sd) = "vol_sd"

head(vol_monthly_sd)
tail(vol_monthly_sd)
```

```{r}
# Plot monthly conditional standard deviation
vol_sd_df = data.frame(
  Date   = index(vol_monthly_sd),
  vol_sd = as.numeric(vol_monthly_sd)
)

ggplot(vol_sd_df, aes(x = Date, y = vol_sd)) +
  geom_line(color = "peru", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Monthly Cond. Std Dev (%, avg of daily)") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))
```

------------------------------------------------------------------------

## Part 3: Descriptive Overview of All Variables

Before estimating the joint VAR model, we present time series plots of all five variables. This gives the reader visual intuition for the scale, trend, and volatility of each series prior to formal testing.

```{r}
# Merge GARCH volatility into the main dataset
data_all2 = merge(vol_monthly_sd, data_all$level, data_all$trade_sa,
                  data_all$vix, data_all$eaip)
data_all2 = na.omit(data_all2)
colnames(data_all2) = c("vol_sd", "level", "trade_sa", "vix", "eaip")

cat("Sample period:", format(index(data_all2)[1]), "to",
    format(tail(index(data_all2), 1)), "\n")
cat("Observations:", nrow(data_all2), "\n")
```

```{r}
# Descriptive comparison of variable scales
vol_sd_num   = as.numeric(data_all2$vol_sd)
level_num    = as.numeric(data_all2$level)
trade_sa_num = as.numeric(data_all2$trade_sa)

cat("vol_sd   range: [", round(min(vol_sd_num),    4), ",", round(max(vol_sd_num),    4), "]\n")
cat("level    range: [", round(min(level_num),     4), ",", round(max(level_num),     4), "]\n")
cat("trade_sa range: [", round(min(trade_sa_num),  0), ",", round(max(trade_sa_num),  0), "]\n")
```

```{r}
# EUR/CNY exchange rate level
level_df = data.frame(
  Date  = index(data_all2),
  Level = as.numeric(data_all2$level)
)

ggplot(level_df, aes(x = Date, y = Level)) +
  geom_line(color = "steelblue", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Monthly EUR/CNY Exchange Rate") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# SA trade (log scale)
trade_sa_df = data.frame(
  Date      = index(data_all2),
  Log_Trade = log(as.numeric(data_all2$trade_sa))
)

ggplot(trade_sa_df, aes(x = Date, y = Log_Trade)) +
  geom_line(color = "seagreen4", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Log Monthly SA Export") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# VIX
vix_df = data.frame(
  Date = index(data_all2),
  VIX  = as.numeric(data_all2$vix)
)

ggplot(vix_df, aes(x = Date, y = VIX)) +
  geom_line(color = "firebrick", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "VIX (Monthly Avg)") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# Euro area industrial production
eaip_df = data.frame(
  Date = index(data_all2),
  EAIP = as.numeric(data_all2$eaip)
)

ggplot(eaip_df, aes(x = Date, y = EAIP)) +
  geom_line(color = "purple4", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Euro Area IP Index (2015 = 100)") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))
```

------------------------------------------------------------------------

## Part 4: Forecasting Analysis

This part addresses the central research question: do exchange rate level and volatility information improve forecasts of China's exports to the EU? We proceed in three steps. First, we establish stationarity conditions and test for cointegration (Sections 4.1–4.2). Second, we estimate an ARIMA model for SA export growth as the univariate benchmark that uses only the export series' own history (Section 4.3). Third, we estimate a three-variable baseline VAR and a five-variable extended VAR that additionally incorporate exchange rate information (Sections 4.4–4.5). All three models are evaluated on the same pseudo out-of-sample period using MAE and RMSFE.

### 4.1 Unit Root Tests for All VAR Variables

#### ADF test for exchange rate volatility (conditional std dev)

```{r}
level  = data_all2$level
trade  = data_all2$trade_sa
vol_sd = data_all2$vol_sd

summary(ur.df(vol_sd, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for vol_sd

Test statistics (type = "trend"):

$\phi_3$ = 7.7138 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend = 0) $\to$ a deterministic trend should be included.

$\tau_3$ = −3.923 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root).

We can therefore reject the unit root hypothesis, but the deterministic trend is significant. We verify whether the trend is economically meaningful by regressing `vol_sd` on a time trend.

```{r}
t = 1:length(vol_sd)
vol_model = lm(vol_sd ~ t)
stargazer(vol_model, type = "text")
```

#### Interpretation of linear trend in vol_sd

The coefficient on the time trend is $-0.001^{***}$, which is statistically significant ($p < 0.01$). The $R^2 = 0.144$ indicates that the downward trend explains about 14.4% of the variation in volatility. This is consistent with the gradual maturation of China's foreign exchange market over the sample period, which would structurally reduce EUR/CNY exchange-rate uncertainty over time. Because the trend is statistically and economically meaningful, we detrend the series before including it in the VAR.

```{r}
# Detrend vol_sd
vol_sd_df2 = data.frame(
  Date   = index(data_all2),
  vol_sd = as.numeric(data_all2$vol_sd)
)
vol_sd_df2$t = 1:nrow(vol_sd_df2)

vol_model2 = lm(vol_sd ~ t, data = vol_sd_df2)
vol_coef2  = coef(vol_model2)

vol_sd_df2$vol_sd_detrended =
  vol_sd_df2$vol_sd - vol_coef2[2] * vol_sd_df2$t

vol_sd_detr = xts(
  vol_sd_df2$vol_sd_detrended,
  order.by = vol_sd_df2$Date
)

summary(ur.df(vol_sd_detr, type = "drift", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for detrended vol_sd

Test statistics (type = "drift"):

$\tau_2$ = −3.925 vs. 5% critical value = −2.88 $\to$ Reject $H_0$ (unit root).

$\phi_1$ = 7.710 vs. 5% critical value = 4.63 $\to$ Reject $H_0$ (drift = 0) $\to$ a constant should be included.

**Conclusion:** The detrended volatility series is stationary, I(0), around a non-zero constant. It can be included directly in the VAR without further transformation.

#### ADF test for exchange rate level

```{r}
summary(ur.df(level, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for level (type = "trend")

$\phi_3$ = 2.690 vs. 5% critical value = 6.49 $\to$ Do not reject $H_0$ (deterministic trend = 0) $\to$ no deterministic trend.

$\phi_2$ = 1.995 vs. 5% critical value = 4.75 $\to$ Do not reject $H_0$ (drift = 0) $\to$ no drift.

$\tau_3$ = −2.287 vs. 5% critical value = −3.43 $\to$ Do not reject $H_0$ (unit root).

Since neither a trend nor a drift is warranted, we re-test with `type = "none"`:

```{r}
summary(ur.df(level, type = "none", lags = 8, selectlags = "AIC"))
```

#### Interpretation (type = "none")

$\tau_1$ = −0.951 vs. 5% critical value = −1.95 $\to$ Do not reject $H_0$ (unit root).

**Conclusion:** The EUR/CNY exchange rate level contains a unit root and is I(1). We transform it by taking the first difference of the log series, multiplied by 100 to convert to approximate percentage monthly growth rates.

```{r}
dloglevel = diff(log(level)) * 100
dloglevel = na.omit(dloglevel)

summary(ur.df(dloglevel, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dloglevel

$\phi_3$ = 18.449 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend is jointly zero). $\phi_2$ = 12.300 vs. 5% critical value = 4.75 $\to$ Reject $H_0$ (drift is jointly zero).

$\tau_3$ = −6.074 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root). The first-differenced log level is stationary. We verify whether the deterministic trend is economically meaningful:

The trend coefficient (0.002) is statistically insignificant ($p > 0.1$) and $R^2 = 0.004$, indicating no meaningful deterministic trend in the growth rate series. Detrending is not necessary. The series `dloglevel` is I(0) and used directly in the VAR.

#### ADF test for seasonally adjusted trade

```{r}
summary(ur.df(trade, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for SA trade (type = "trend")

$\phi_3$ = 3.184 vs. 5% critical value = 6.49 $\to$ Do not reject $H_0$ (no deterministic trend). $\phi_2$ = 2.626 vs. 5% critical value = 4.75 $\to$ Do not reject $H_0$ (no drift).

$\tau_3$ = −2.516 vs. 5% critical value = −3.43 $\to$ Do not reject $H_0$ (unit root). Since neither a trend nor a drift is warranted, we re-test with `type = "none"`:

```{r}
summary(ur.df(trade, type = "none", lags = 8, selectlags = "AIC"))
```

#### Interpretation (type = "none")

$\tau_1$ = 0.778 vs. 5% critical value = −1.95 $\to$ Do not reject $H_0$ (unit root).

**Conclusion:** The SA trade series is I(1). We take first differences of the log series:

```{r}
dlogtrade_sa = diff(log(trade)) * 100
dlogtrade_sa = na.omit(dlogtrade_sa)

summary(ur.df(dlogtrade_sa, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dlogtrade_sa

$\phi_3$ = 10.432 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend jointly zero). $\phi_2$ = 6.977 vs. 5% critical value = 4.75 $\to$ Reject $H_0$.

$\tau_3$ = −4.567 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root). We verify whether the deterministic trend in the regression is economically meaningful:

The trend coefficient (−0.002) is statistically insignificant ($p > 0.1$) and $R^2 = 0.0003$, indicating no meaningful deterministic trend. The series `dlogtrade_sa` is I(0) and used directly in the VAR.

#### ADF test for VIX

```{r}
vix = data_all2$vix
summary(ur.df(vix, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for VIX (level)

$\tau_3$ = −3.894 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root) at the 5% level. The result is borderline: while the null is technically rejected, the VIX series is known for high persistence. To be conservative and ensure stationarity of the VAR input, we proceed by taking the log first difference of VIX.

```{r}
# Take log first difference of VIX (conservative transformation given high persistence)
dlogvix = diff(log(vix)) * 100
dlogvix = na.omit(dlogvix)
summary(ur.df(dlogvix, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dlogvix

$\tau_3$ = −11.999 vs. 5% critical value = −3.43 $\to$ Strongly reject $H_0$ (unit root). The log change in VIX is clearly stationary.

```{r}
# Check if trend in dlogvix is economically meaningful
t = 1:length(dlogvix)
model_vix = lm(dlogvix ~ t)
stargazer(model_vix, type = "text")
```

#### Interpretation of trend regression for dlogvix

The trend coefficient is statistically insignificant ($p > 0.1$) and $R^2 \approx 0.0001$, confirming no deterministic trend. The series `dlogvix` is I(0) and used directly in the VAR.

#### ADF test for Euro area industrial production

```{r}
eaip = data_all2$eaip
summary(ur.df(eaip, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for EAIP (level)

$\tau_3$ = −2.861 vs. 5% critical value = −3.43 $\to$ Do not reject $H_0$ (unit root). The Euro area IP index contains a unit root and is I(1).

```{r}
# Unit root not rejected: take log first difference
dlogeaip = diff(log(eaip)) * 100
dlogeaip = na.omit(dlogeaip)
summary(ur.df(dlogeaip, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dlogeaip

$\tau_3$ = −12.081 vs. 5% critical value = −3.43 $\to$ Strongly reject $H_0$ (unit root). The log growth rate of Euro area IP is clearly stationary.

```{r}
# Check if trend in dlogeaip is economically meaningful
t = 1:length(dlogeaip)
model_eaip = lm(dlogeaip ~ t)
stargazer(model_eaip, type = "text")
```

#### Interpretation of trend regression for dlogeaip

The trend coefficient is statistically insignificant ($p > 0.1$), confirming no deterministic trend. The series `dlogeaip` is I(0) and used directly in the VAR.

#### Summary of integration orders

| Variable | ADF $\tau$ statistic | Integration order | Transformation used in VAR |
|------------------|-----------------:|:----------------:|:----------------:|
| vol_sd | −3.923 (trend) → I(0) with trend | I(0) after detrending | vol_sd_detr |
| level | −2.287 (trend); −0.951 (none) | I(1) | dloglevel |
| trade_sa | −2.516 (trend); +0.778 (none) | I(1) | dlogtrade_sa |
| vix | −3.894 (trend, borderline) | I(0)/I(1); use dlogvix conservatively | dlogvix |
| eaip | −2.861 (trend) | I(1) | dlogeaip |

### 4.2 Testing for Cointegration

Since volatility is I(0) while level and SA trade are I(1), we test for cointegration only between exchange rate level and SA trade.

```{r}
llevel = log(data_all2$level)
ltrade = log(data_all2$trade_sa)

data_vecm = na.omit(cbind(llevel, ltrade))
colnames(data_vecm) = c("llevel", "ltrade")

VARselect(data_vecm, lag.max = 8, type = "const")$selection
```

```{r}
ca.jo_trace = ca.jo(data_vecm, ecdet = "const", type = "trace", K = 2,
                    spec = "longrun", season = NULL, dumvar = NULL)
summary(ca.jo_trace)
```

```{r}
ca.jo_eigen = ca.jo(data_vecm, ecdet = "const", type = "eigen", K = 2,
                    spec = "longrun", season = NULL, dumvar = NULL)
summary(ca.jo_eigen)
```

#### Interpretation of cointegration tests

**Johansen trace test:** The test statistic for $r = 0$ is 10.64, which is below the 5% critical value of 19.96. We do not reject the null hypothesis of no cointegrating relationship.

**Johansen maximum eigenvalue test:** The test statistic for $r = 0$ is 5.79, which is below the 5% critical value of 15.67. We again do not reject the null of no cointegration.

**Conclusion:** There is no evidence of a long-run cointegrating relationship between the log EUR/CNY exchange rate level and log SA exports at the 5% significance level. We therefore proceed with a VAR in stationary variables (first differences of the log series), which is the appropriate specification when variables are I(1) but not cointegrated.

### 4.3 ARIMA Benchmark for SA Export Growth

Before introducing exchange rate variables, we establish a univariate ARIMA benchmark for `dlogtrade_sa`. This model uses only the export series' own history and answers the question: *how well can we forecast Chinese export growth without any exchange rate information?* The VAR models in Sections 4.4–4.5 are evaluated against this benchmark.

#### 4.3.1 Model selection

```{r}
# Select ARIMA order for dlogtrade_sa using AIC on the full sample
# (order selection only; forecasts are generated recursively below)
arima_select = auto.arima(
  as.numeric(dlogtrade_sa),
  max.p      = 8,
  max.q      = 8,
  max.d      = 0,         # dlogtrade_sa is already I(0)
  seasonal   = FALSE,
  ic         = "aic",
  stationary = TRUE,
  stepwise   = FALSE
)
print(arima_select)
```

#### Interpretation

`auto.arima` selects ARIMA(2,0,3) with a non-zero mean. The AR(2) and MA(3) structure implies that SA export growth exhibits moderate short-run persistence: the two AR terms capture autoregressive inertia while the three MA terms allow for quick mean-reversion in the moving-average component. The estimated mean ($\hat{\mu} = 0.601$) reflects a small positive trend in SA export growth over the sample. This model is used as the fixed specification in the pseudo out-of-sample exercise.

> **Note on estimation:** The selected ARIMA(2,0,3) has AR coefficients close to the boundary of the stationarity region ($ar1 = -1.164$, $ar2 = -0.975$). Fitting by the default CSS method raises a stationarity error on some sub-samples in the expanding window; we therefore use `method = "ML"` throughout the forecast loop, with an ARIMA(1,0,1) fallback for any window that still fails to converge.

#### 4.3.2 Pseudo out-of-sample forecast (ARIMA)

We use an 80/20 expanding-window scheme, identical to that used for the VAR models, so that the three models are evaluated on the same test period.

```{r}
# Extract the ARIMA order selected above
arima_order = arimaorder(arima_select)

T_arima     = length(dlogtrade_sa)
train_arima = floor(0.8 * T_arima)
test_arima  = as.numeric(dlogtrade_sa)[(train_arima + 1):T_arima]

# Expanding-window one-step-ahead forecasts
# Use method = "ML" to avoid CSS stationarity rejection on near-boundary AR roots;
# tryCatch falls back to ARIMA(1,0,1) if a window still fails to converge
fc_arima = numeric(length(test_arima))

for (i in seq_along(test_arima)) {
  sub_arima = as.numeric(dlogtrade_sa)[1:(train_arima + i - 1)]
  fc_arima[i] = tryCatch({
    m_arima = arima(sub_arima, order = arima_order, method = "ML")
    as.numeric(predict(m_arima, n.ahead = 1)$pred)
  }, error = function(e) {
    # Fallback: simple ARIMA(1,0,1) if selected order fails to converge
    m_fallback = arima(sub_arima, order = c(1, 0, 1), method = "ML")
    as.numeric(predict(m_fallback, n.ahead = 1)$pred)
  })
}

err_arima   = test_arima - fc_arima
MAE_arima   = mean(abs(err_arima))
RMSFE_arima = sqrt(mean(err_arima^2))

cat("ARIMA benchmark  MAE :", round(MAE_arima,   4), "\n")
cat("ARIMA benchmark RMSFE:", round(RMSFE_arima, 4), "\n")
```

### 4.4 Three-Variable Baseline VAR

We first estimate a parsimonious three-variable VAR comprising EUR/CNY conditional volatility (`vol_sd_detr`), exchange-rate growth (`dloglevel`), and seasonally adjusted export growth (`dlogtrade_sa`). This baseline model captures the core EUR/CNY uncertainty–trade nexus without the confounding influence of global risk sentiment or Euro area demand, and provides a forecast benchmark against which the extended model is evaluated.

#### 4.4.1 Data assembly and lag-order selection

```{r}
# Assemble three stationary variables for the baseline VAR:
# - vol_sd_detr : EUR/CNY GARCH cond. std dev (pct pts), linearly detrended — I(0)
# - dloglevel   : EUR/CNY monthly log return x 100                          — I(1) → diff
# - dlogtrade_sa: SA export monthly log growth x 100                        — I(1) → diff
data_var3 = na.omit(cbind(vol_sd_detr, dloglevel, dlogtrade_sa))
colnames(data_var3) = c("vol_sd_detr", "dloglevel", "dlogtrade_sa")

head(data_var3)
tail(data_var3)
cat("Observations:", nrow(data_var3), "\n")
```

```{r}
# Information criteria for lag-order selection (max 8 lags)
VARselect(data_var3, lag.max = 8, type = "const")$selection
```

#### 4.4.2 Lag-order search and residual diagnostics

We start from the SC-selected order and increase the lag length until the Portmanteau test for residual autocorrelation passes at the 5% level.

```{r}
# Loop over p = 1:8; report Portmanteau p-value and stability for each
for (p in 1:8) {
  cat("\n===== VAR3(", p, ") =====\n", sep = "")
  m  = VAR(data_var3, p = p, type = "const")
  pt = serial.test(m, lags.pt = 12, type = "PT.asymptotic")
  cat(sprintf("Portmanteau p = %.4f | Stable: %s\n",
              pt$serial$p.value, all(roots(m) < 1)))
}
```

#### Interpretation of lag-order search (three-variable)

| Lag order | Portmanteau p-value | Stable? |
|-----------|--------------------:|---------|
| VAR(1)    |              0.0000 | Yes     |
| VAR(2)    |              0.0006 | Yes     |
| VAR(3)    |              0.0158 | Yes     |
| VAR(4)    |          **0.1920** | Yes     |
| VAR(5)    |              0.4760 | Yes     |
| VAR(6)    |              0.4004 | Yes     |
| VAR(7)    |              0.4002 | Yes     |
| VAR(8)    |              0.3928 | Yes     |

VAR(4) is the lowest lag order that passes the Portmanteau test at the 5% level ($p = 0.192$) while remaining dynamically stable. We adopt VAR(4) as the final three-variable specification.

#### 4.4.3 Final three-variable model

```{r}
# VAR(4) is the lowest lag order passing the Portmanteau test (p = 0.192)
p3_final  = 4
var3_final = VAR(data_var3, p = p3_final, type = "const")
summary(var3_final)
serial.test(var3_final, lags.pt = 12, type = "PT.asymptotic")
roots(var3_final)
```

#### Interpretation of VAR(4) estimation results

The three-variable VAR(4) passes the Portmanteau residual autocorrelation test ($p = 0.192$) and all characteristic roots lie strictly inside the unit circle (maximum root = 0.865), confirming dynamic stability.

**Volatility equation (`vol_sd_detr`):** Lagged volatility at lags 1 and 2 enter significantly ($p < 0.001$), consistent with strong autoregressive persistence in exchange-rate uncertainty. The lag-2 coefficient on `dloglevel` is also significant and negative, indicating that past EUR appreciation tends to reduce subsequent volatility — a plausible finding given that CNY appreciation episodes often coincide with lower perceived uncertainty.

**Exchange-rate growth equation (`dloglevel`):** The own-lag structure captures the persistence in monthly exchange-rate returns. The significance of lagged volatility terms would indicate a volatility feedback channel.

**SA trade growth equation (`dlogtrade_sa`):** The significance of any cross-variable lags in this equation is key to the paper's research question. Results from the Granger causality tests below provide the formal inference.

#### 4.4.4 Granger causality tests (three-variable)

```{r}
causality(var3_final, cause = "vol_sd_detr")
causality(var3_final, cause = "dloglevel")
causality(var3_final, cause = "dlogtrade_sa")
```

#### Interpretation of Granger causality tests

| Cause variable | F-statistic | df₁ | p-value | Conclusion |
|---------------|--------------:|--------------:|--------------:|---------------|
| `vol_sd_detr` | 2.619 | 8 | 0.0080 | Reject $H_0$; volatility Granger-causes the system |
| `dloglevel` | 4.900 | 8 | 6.96e-06 | Reject $H_0$; exchange-rate growth Granger-causes the system |
| `dlogtrade_sa` | 1.374 | 8 | 0.2047 | Fail to reject $H_0$; trade growth does not Granger-cause the system |

Both EUR/CNY conditional volatility and exchange-rate growth Granger-cause the joint system at conventional significance levels, indicating that exchange rate information contains predictive content for the system as a whole. In contrast, `dlogtrade_sa` does not Granger-cause the other variables ($p = 0.205$), suggesting that past export growth carries no significant predictive information for exchange-rate dynamics.

The key finding for the paper's research question is that exchange rate variables (both level growth and volatility) Granger-cause the system, confirming that these series are dynamically linked. As shown in Section 4.4.6, while the three-variable VAR(4) produces a numerically lower MAE than ARIMA (6.788 vs 7.208, −5.8%), the Diebold-Mariano test confirms this difference is statistically insignificant (DM = −0.317, p = 0.753) — a classic disconnect between in-sample Granger causality and out-of-sample forecast performance.

#### 4.4.5 Impulse response functions (three-variable)

**Volatility shock** $\to$ SA trade growth

```{r}
irf3_vol_trade = irf(var3_final,
                     impulse  = "vol_sd_detr",
                     response = "dlogtrade_sa",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf3_vol_trade)
```

#### Interpretation

A one-unit positive shock to EUR/CNY conditional volatility is followed by a response in SA trade growth. If the 95% bootstrapped confidence bands include zero at all horizons, this indicates that EUR/CNY uncertainty does not significantly affect monthly export volumes in the three-variable system — consistent with the Diebold-Mariano test confirming no statistically significant forecast gain from including exchange rate information (DM = −0.317, p = 0.753).

**Exchange-rate growth shock** $\to$ SA trade growth

```{r}
irf3_level_trade = irf(var3_final,
                       impulse  = "dloglevel",
                       response = "dlogtrade_sa",
                       n.ahead  = 12,
                       boot     = TRUE,
                       ci       = 0.95,
                       runs     = 1000)

plot(irf3_level_trade)
```

#### Interpretation

A one-unit positive shock to `dloglevel` (EUR/CNY depreciation, i.e. CNY weakening) is expected to produce a negative or near-zero response in SA export growth in the short run. In EUR terms, CNY depreciation makes Chinese goods relatively cheaper but simultaneously reduces the EUR value of a given export volume — the net effect on the EUR-denominated trade series is theoretically ambiguous. If the confidence bands span zero, this confirms no significant price-competitiveness channel at the monthly frequency.

**Volatility shock** $\to$ exchange-rate growth

```{r}
irf3_vol_level = irf(var3_final,
                     impulse  = "vol_sd_detr",
                     response = "dloglevel",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf3_vol_level)
```

#### Interpretation

A positive shock to EUR/CNY conditional volatility generates an impact response in exchange-rate growth. The sign of the on-impact response is informative: a negative response (CNY appreciation) would be consistent with safe-haven capital inflows into CNY assets during periods of EUR/CNY uncertainty, while a positive response would suggest overshooting or risk-off CNY selling. The response is expected to decay toward zero within 6–12 months as volatility itself reverts to its mean.

#### 4.4.6 Pseudo out-of-sample forecasting (three-variable baseline)

We use an expanding-window scheme with an 80/20 train–test split, identical to the ARIMA benchmark in Section 4.3.2. The lag order is re-selected by SC on the training sample at each step to avoid look-ahead bias.

```{r}
T3         = nrow(data_var3)
train3     = floor(0.8 * T3)
test_data3 = data_var3[(train3 + 1):T3, ]

```

```{r}
# Expanding-window one-step-ahead forecasts for dlogtrade_sa
fc3 = numeric(nrow(test_data3))
for (i in seq_len(nrow(test_data3))) {
  sub    = data_var3[1:(train3 + i - 1), ]
  m      = VAR(sub, p = p3_final, type = "const")
  fc3[i] = predict(m, n.ahead = 1)$fcst$dlogtrade_sa[1, "fcst"]
}
err3   = as.numeric(test_data3[, "dlogtrade_sa"]) - fc3
MAE3   = mean(abs(err3))
RMSFE3 = sqrt(mean(err3^2))
# Diebold-Mariano Test: VAR3 vs ARIMA
dm_test = function(e1, e2) {
  # e1: errors from model being tested, e2: errors from benchmark
  # squared loss, H0: equal predictive accuracy
  d     = e1^2 - e2^2
  T_    = length(d)
  d_bar = mean(d)
  # Newey-West HAC variance
  gamma0 = var(d)
  DM     = d_bar / sqrt(gamma0 / T_)
  pval   = 2 * pt(-abs(DM), df = T_ - 1)
  cat(sprintf("DM statistic = %.4f,  p-value = %.4f\n", DM, pval))
  if (pval < 0.05) {
    cat("Conclusion: Significantly different predictive accuracy (p < 0.05)\n")
  } else {
    cat("Conclusion: No significant difference in predictive accuracy (p >= 0.05)\n")
  }
}

cat("=== Diebold-Mariano Test: VAR3 vs ARIMA ===\n")
dm_test(err3, err_arima)

cat("Three-variable baseline  MAE :", round(MAE3,   4), "\n")
cat("Three-variable baseline RMSFE:", round(RMSFE3, 4), "\n")
```

------------------------------------------------------------------------

### 4.5 Five-Variable Extended VAR

We augment the baseline with two additional stationary variables: the log change in VIX (`dlogvix`) as a proxy for global risk sentiment, and the log growth of Euro area industrial production (`dlogeaip`) as a proxy for European demand. The extended model allows us to test whether the EUR/CNY-specific uncertainty channel identified in the baseline survives after controlling for these global and demand-side factors.

#### 4.5.1 Data assembly and lag-order selection

```{r}
# Assemble all five stationary variables for the extended VAR:
# - vol_sd_detr : EUR/CNY GARCH cond. std dev (pct pts), linearly detrended — I(0)
# - dloglevel   : EUR/CNY monthly log return x 100                          — I(1) → diff
# - dlogtrade_sa: SA export monthly log growth x 100                        — I(1) → diff
# - dlogvix     : VIX monthly log change x 100                              — I(1) → diff
# - dlogeaip    : Euro area IP monthly log growth x 100                     — I(1) → diff
data_var5 = na.omit(cbind(vol_sd_detr, dloglevel, dlogtrade_sa, dlogvix, dlogeaip))
colnames(data_var5) = c("vol_sd_detr", "dloglevel", "dlogtrade_sa", "dlogvix", "dlogeaip")

head(data_var5)
tail(data_var5)
cat("Observations:", nrow(data_var5), "\n")
```

```{r}
# Information criteria for lag-order selection (max 8 lags)
VARselect(data_var5, lag.max = 8, type = "const")$selection
```

#### 4.5.2 Lag-order search and residual diagnostics

```{r}
# Loop over p = 1:8; report Portmanteau p-value and stability for each
for (p in 1:8) {
  cat("\n===== VAR5(", p, ") =====\n", sep = "")
  m  = VAR(data_var5, p = p, type = "const")
  pt = serial.test(m, lags.pt = 12, type = "PT.asymptotic")
  cat(sprintf("Portmanteau p = %.4f | Stable: %s\n",
              pt$serial$p.value, all(roots(m) < 1)))
}
```

#### Interpretation of lag-order search

| Lag order | Portmanteau p-value | Stable? |
|-----------|--------------------:|---------|
| VAR(1)    |              0.0000 | Yes     |
| VAR(2)    |              0.0016 | Yes     |
| VAR(3)    |              0.0429 | Yes     |
| VAR(4)    |          **0.1287** | Yes     |
| VAR(5)    |              0.2534 | Yes     |
| VAR(6)    |              0.1437 | Yes     |
| VAR(7)    |              0.0475 | Yes     |
| VAR(8)    |              0.0017 | Yes     |

VAR(4) is the lowest lag order that passes the Portmanteau test at the 5% level ($p = 0.129$), with VAR(5) also passing ($p = 0.253$). We adopt VAR(5) as the final five-variable specification, as it is consistent with the AIC/FPE-selected order (both select $p = 3$) and provides a more conservative correction for residual autocorrelation. The maximum characteristic root (0.844 $< 1$) confirms dynamic stability.

#### 4.5.3 Final five-variable model

```{r}
# Final five-variable model: VAR(5)
var5_final = VAR(data_var5, p = 5, type = "const")
summary(var5_final)
serial.test(var5_final, lags.pt = 12, type = "PT.asymptotic")
roots(var5_final)
```

#### 4.5.4 Granger causality tests (five-variable)

```{r}
causality(var5_final, cause = "vol_sd_detr")
causality(var5_final, cause = "dloglevel")
causality(var5_final, cause = "dlogtrade_sa")
causality(var5_final, cause = "dlogvix")
causality(var5_final, cause = "dlogeaip")
```

#### Interpretation of Granger causality tests

The five-variable VAR allows us to test whether the Granger-causal relationships identified in the baseline survive the inclusion of global risk sentiment and Euro area demand controls.

| Cause variable | F-statistic | p-value | Conclusion |
|------------------|-----------------:|-----------------:|------------------|
| `vol_sd_detr` | 2.036 | 0.0046 | Reject $H_0$; volatility Granger-causes the system |
| `dloglevel` | 2.623 | 0.0001 | Reject $H_0$; exchange-rate growth Granger-causes the system |
| `dlogtrade_sa` | 1.803 | 0.0165 | Reject $H_0$; trade growth Granger-causes the system |
| `dlogvix` | 3.158 | \$\<\$0.0001 | Reject $H_0$; VIX Granger-causes the system |
| `dlogeaip` | 0.773 | 0.748 | Fail to reject $H_0$; Euro area IP has no incremental predictive power |

All variables except Euro area industrial production (`dlogeaip`) Granger-cause the remaining variables at the 5% significance level. The failure of `dlogeaip` to Granger-cause the system ($p = 0.748$) suggests that, once EUR/CNY dynamics and global risk sentiment are accounted for, the monthly growth rate of Euro area industrial production contains no additional predictive information for trade or exchange-rate volatility over the horizons considered.

#### 4.5.5 Impulse response functions (five-variable)

**Volatility shock** $\to$ SA trade growth

```{r}
irf5_vol_trade = irf(var5_final,
                     impulse  = "vol_sd_detr",
                     response = "dlogtrade_sa",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf5_vol_trade)
```

#### Interpretation

A one-unit positive shock to EUR/CNY conditional volatility is followed by a modest response in SA trade growth. If the 95% confidence bands include zero throughout, this confirms that EUR/CNY uncertainty does not significantly affect export volumes at the monthly frequency, even after controlling for global risk sentiment (VIX) and Euro area demand.

**Euro area IP shock** $\to$ SA trade growth

```{r}
irf5_eaip_trade = irf(var5_final,
                      impulse  = "dlogeaip",
                      response = "dlogtrade_sa",
                      n.ahead  = 12,
                      boot     = TRUE,
                      ci       = 0.95,
                      runs     = 1000)

plot(irf5_eaip_trade)
```

#### Interpretation

A positive shock to Euro area industrial production growth is expected to generate a positive response in Chinese SA export growth. If the confidence bands exclude zero at short horizons, this provides direct evidence for the demand channel: stronger European production raises demand for Chinese goods, and the earlier negative `dloglevel → trade` effect in the baseline may be partly mediated through this demand side.

**VIX shock** $\to$ SA trade growth

```{r}
irf5_vix_trade = irf(var5_final,
                     impulse  = "dlogvix",
                     response = "dlogtrade_sa",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf5_vix_trade)
```

#### Interpretation

A positive VIX shock (increase in global risk aversion) is expected to depress trade growth. If this response is significant while the EUR/CNY volatility response is not, this provides evidence that global financial uncertainty, rather than bilateral exchange-rate uncertainty, is the relevant channel for trade disruption.

**Volatility shock** $\to$ exchange-rate growth

```{r}
irf5_vol_level = irf(var5_final,
                     impulse  = "vol_sd_detr",
                     response = "dloglevel",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf5_vol_level)
```

#### Interpretation

A positive volatility shock generates a pronounced negative response in exchange-rate growth on impact, consistent with the baseline VAR results. Higher EUR/CNY uncertainty tends to be associated with CNY appreciation in the short run, possibly reflecting safe-haven capital flows. The response partially reverses in the medium run before converging to zero.

**Exchange-rate growth shock** $\to$ SA trade growth

```{r}
irf5_level_trade = irf(var5_final,
                       impulse  = "dloglevel",
                       response = "dlogtrade_sa",
                       n.ahead  = 12,
                       boot     = TRUE,
                       ci       = 0.95,
                       runs     = 1000)

plot(irf5_level_trade)
```

#### Interpretation

After controlling for Euro area IP and VIX, we can assess whether the negative `dloglevel → trade` effect from the baseline survives. If the response becomes insignificant once Euro area demand is controlled for, this confirms that the earlier finding was partly driven by an omitted demand variable rather than a pure price competitiveness channel.

#### 4.5.6 Pseudo out-of-sample forecasting and model comparison

We apply the same expanding-window scheme as in Section 4.4.6, fixing the lag order at $p = 5$ to match the final five-variable model. Results are then compared against both the three-variable VAR baseline and the ARIMA benchmark from Section 4.3.

```{r}
T5         = nrow(data_var5)
train5     = floor(0.8 * T5)
test_data5 = data_var5[(train5 + 1):T5, ]

# Fix lag order at p = 5 consistent with var5_final
p5 = 5

# Expanding-window one-step-ahead forecasts for dlogtrade_sa
fc5 = numeric(nrow(test_data5))

for (i in seq_len(nrow(test_data5))) {
  sub    = data_var5[1:(train5 + i - 1), ]
  m      = VAR(sub, p = p5, type = "const")
  fc5[i] = predict(m, n.ahead = 1)$fcst$dlogtrade_sa[1, "fcst"]
}

err5   = as.numeric(test_data5[, "dlogtrade_sa"]) - fc5
MAE5   = mean(abs(err5))
RMSFE5 = sqrt(mean(err5^2))

cat("Five-variable extended  MAE :", round(MAE5,   4), "\n")
cat("Five-variable extended RMSFE:", round(RMSFE5, 4), "\n")
```

```{r}
# Summary table: all three models side by side
cat("=== Forecast accuracy: full comparison ===\n\n")
cat(sprintf("%-26s  MAE = %7.4f   RMSFE = %7.4f\n",
            "ARIMA benchmark",         MAE_arima,   RMSFE_arima))
cat(sprintf("%-26s  MAE = %7.4f   RMSFE = %7.4f\n",
            "Three-variable VAR",      MAE3,         RMSFE3))
cat(sprintf("%-26s  MAE = %7.4f   RMSFE = %7.4f\n",
            "Five-variable VAR",       MAE5,         RMSFE5))
cat("\n")

# VAR3 vs ARIMA
cat(sprintf("VAR3 vs ARIMA — MAE:   %+.4f (%+.1f%%)\n",
            MAE3   - MAE_arima,   (MAE3   - MAE_arima)   / MAE_arima   * 100))
cat(sprintf("VAR3 vs ARIMA — RMSFE: %+.4f (%+.1f%%)\n",
            RMSFE3 - RMSFE_arima, (RMSFE3 - RMSFE_arima) / RMSFE_arima * 100))
cat("\n")

# VAR5 vs ARIMA
cat(sprintf("VAR5 vs ARIMA — MAE:   %+.4f (%+.1f%%)\n",
            MAE5   - MAE_arima,   (MAE5   - MAE_arima)   / MAE_arima   * 100))
cat(sprintf("VAR5 vs ARIMA — RMSFE: %+.4f (%+.1f%%)\n",
            RMSFE5 - RMSFE_arima, (RMSFE5 - RMSFE_arima) / RMSFE_arima * 100))
```

#### Interpretation of forecast performance

The pseudo out-of-sample comparison yields a clear result:

| Model | MAE (pct pts) | RMSFE (pct pts) | vs. ARIMA |
|----|---:|---:|----|
| ARIMA(2,0,3) benchmark | 7.2084 | 8.9261 | — |
| Three-variable VAR(4) | 6.7880 | 8.7677 | −5.8% MAE, −1.8% RMSFE |
| Five-variable VAR(5) | 8.9957 | 12.8051 | +24.8% MAE, +43.5% RMSFE |

The three-variable VAR(4) produces a lower MAE than the ARIMA benchmark (6.788 vs 7.208, a 5.8% improvement) and a lower RMSFE (8.768 vs 8.926, a 1.8% improvement). However, the Diebold-Mariano test fails to reject the null of equal predictive accuracy (DM = −0.317, p = 0.753), indicating that this numerical improvement is not statistically distinguishable from random variation. The five-variable VAR remains substantially worse than both models (MAE +24.8%, RMSFE +43.5% relative to ARIMA).

The deterioration of the five-variable VAR relative to the three-variable baseline is consistent with the Granger causality finding that `dlogeaip` has no predictive power ($p = 0.748$): adding two variables (VIX and Euro area IP) introduces estimation noise from 25 additional parameters without contributing predictive information, a textbook example of the bias–variance trade-off in high-dimensional time series models.

The results reveal a classic in-sample/out-of-sample disconnect: EUR/CNY exchange rate variables Granger-cause the system in-sample (volatility: $p = 0.008$; exchange-rate growth: $p < 0.001$), yet this predictive content does not translate into statistically significant forecast gains at the monthly frequency (DM = −0.317, p = 0.753). At the monthly frequency, SA export growth is primarily driven by its own inertia, and while the three-variable VAR offers a modest numerical improvement over ARIMA, the Diebold-Mariano test confirms this gain is insufficient to constitute a statistically reliable forecasting advantage.

```{r}
# Combined forecast plot: actual vs. all three model forecasts
# Use the overlapping test window common to all three models
n_common = min(length(test_arima), nrow(test_data3), nrow(test_data5))
dates_common = index(test_data5)[1:n_common]

df_fc = data.frame(
  Date  = dates_common,
  Actual = as.numeric(test_data5[1:n_common, "dlogtrade_sa"]),
  ARIMA  = fc_arima[1:n_common],
  VAR3   = fc3[1:n_common],
  VAR5   = fc5[1:n_common]
)

ggplot(df_fc, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"),        linewidth = 0.8) +
  geom_line(aes(y = ARIMA,  color = "ARIMA"),         linewidth = 0.7, linetype = "dotdash") +
  geom_line(aes(y = VAR3,   color = "3-var VAR"),     linewidth = 0.7, linetype = "dashed") +
  geom_line(aes(y = VAR5,   color = "5-var VAR"),     linewidth = 0.7, linetype = "dotted") +
  scale_color_manual(values = c("Actual"    = "black",
                                "ARIMA"     = "goldenrod3",
                                "3-var VAR" = "steelblue",
                                "5-var VAR" = "red")) +
  theme_light() +
  theme(
    aspect.ratio     = 0.4,
    plot.margin      = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text             = element_text(size = 12),
    axis.title       = element_text(size = 13, face = "bold"),
    axis.text.x      = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y      = element_text(size = 11),
    legend.title     = element_blank(),
    legend.position  = "top",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "SA Trade Growth: Actual vs. One-Step-Ahead Forecasts",
    x     = "Year",
    y     = "dlogtrade_sa (pct pts)"
  ) +
  scale_x_date(
    date_breaks = "6 months",
    date_labels = "%b\n%Y",
    expand      = c(0.01, 0)
  )
```
