---
title: "Exchange_rate"
author: "Yumeng"
date: "2026-04-08"
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
    number_sections: false
---

# 612 Time Series Project

## Exchange Rate Estimation

Package requirements:

```{r setup, message=FALSE, warning=FALSE}
# install.packages("quantmod")
# install.packages("fredr")
# install.packages("ggfortify")
# install.packages('urca')
# install.packages("tseries")
# install.packages("forecast")
# install.packages("dynlm")
# install.packages("stargazer")
# install.packages("pracma")
# install.packages("dLagM")
# install.packages("gets")
# install.packages("car")
# install.packages("lmtest")
# install.packages("vars")
# install.packages("devtools", type = "binary")
# install.packages("tsDyn")
# install.packages('fGarch', type = "binary")
# No extra package needed: STL adjustment uses base R stats::stl()

options(warn=-1)
```

Library requirements:

```{r}
library(fGarch)
library(xts)
library(urca)
library(stargazer)
library(ggplot2)
library(ggfortify)
library(scales)
library(patchwork)
library(forecast)
library(aTSA)
library(rugarch)
library(lubridate)
library(vars)
# seasonal adjustment uses base R stats::stl() — no extra package needed
```

```{r}
# Import the datasets
fx = read.csv('https://raw.githubusercontent.com/Yumeng20021214/612-project/refs/heads/Yumeng/exchange_rate_final.csv')

# Transfer TIME_PERIOD to Date format
fx$TIME_PERIOD = as.Date(fx$TIME_PERIOD, format = "%m/%d/%Y")

print(head(fx))
print(tail(fx))
print(anyNA(fx))
print(str(fx))
```

```{r}
# Create time series object
fx_xts = xts(fx$OBS_VALUE, order.by = fx$TIME_PERIOD)
colnames(fx_xts) = "OBS_VALUE" 
```

```{r}
dlogfx = diff(log(fx_xts))  # Compute difference in log exchange rate
head(dlogfx)                 # Show first six fx change rate
tail(dlogfx)                 # Show last six fx change rate
rfx = na.omit(dlogfx)       # Remove missing values (from first difference)

summary(ur.df(rfx, type='trend', lags=8, selectlags="AIC"))
```

#### Interpretation of the unit root test

Test statistics ( type = "trend" )

$\phi_3$ = 1379.367 vs. 5% critical value = 6.25 $\to$ Reject $H_0$ (deterministic trend = 0) $\to$ a deterministic trend should be included.

$\tau_3$ = -52.5236 vs. 5% critical value = -3.41 $\to$ Reject $H_0$ (no unit root).

Conclusion: The specification including a deterministic trend is appropriate and provides evidence of no unit root in EUR/CNY exchange rate returns.

```{r}
# Create a deterministic time trend
t = 1:length(rfx)
# Regress return on the time trend
model = lm(rfx ~ t)
# Display regression results in text format
stargazer(model, type = "text")
```

#### Interpretation of the trend regression

Model: $y_t$ = $\alpha$ + $\beta_t$ + $\varepsilon_t$

The estimated trend coefficient ($\beta$) is approximately 0.00000. The coefficient is not statistically significant (p $>$ 0.1). This means there is no significant deterministic trend in the return series.

Conclusion: The EUR/CNY exchange rate returns are stationary with no deterministic trend to remove. The mean of returns remains constant over time, and any fluctuations are random rather than systematic.

```{r}
rfx_df = data.frame(
  Date   = index(rfx),
  Return = as.numeric(rfx)
)

fig <- ggplot(rfx_df, aes(x = Date, y = Return)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "EUR/CNY Exchange Rate Return") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)

fig
```

#### Interpretation of volatility patterns

The figure displays daily EUR/CNY exchange rate returns over a long sample period. The series fluctuates around zero, but the size of the fluctuations changes substantially over time, indicating that volatility is not constant.

2005-2006: Volatility in 2005–2006 was relatively moderate, but it was influenced by important structural changes in China's exchange-rate regime. In 2005, China moved away from the long-standing strict U.S. dollar peg and began shifting toward a more market-based, managed floating system, which increased the role of market forces in RMB pricing.

2008–2009: Volatility rises sharply during the global financial crisis, with much larger positive and negative movements than in earlier years. This is one of the most turbulent periods, and several extreme spikes are visible.

2010–2012: Volatility declines compared with the crisis period, but fluctuations remain more pronounced than in the pre-2008 years, indicating that market uncertainty is still present.

2015–2016: Volatility rose markedly, largely because of China's exchange-rate reform, concerns about China's slowing growth, capital outflow pressures, and uncertainty in global financial markets.

Around 2022: Volatility increased again, mainly because of the conflict between Russia and Ukraine, the resulting energy shock in Europe, and heightened inflation and policy uncertainty.

Around 2025: Volatility after 2025 can be linked to elevated trade-policy uncertainty, tariff shocks, and broader geopolitical tensions.

Overall, the plot suggests that EUR/CNY exchange rate volatility is not constant, but instead tends to increase during periods of economic or financial stress. The visible clustering of large and small movements provides preliminary evidence in favor of using a GARCH-type model to capture time-varying volatility.

### Testing for ARCH Effect

#### Method A: Plot of the squared series

```{r}
rfx2_df = data.frame(
  Date           = index(rfx),
  Squared_Returns = as.numeric(rfx^2)
)

fig <- ggplot(rfx2_df, aes(x = Date, y = Squared_Returns)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Squared Returns") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)

fig
```

#### Interpretation of the plot of squared returns

We observe moderate EUR/CNY return volatility during the early part of the sample. However, volatility increases noticeably during several periods, especially around 2008–2009, 2015–2016, and 2025, where large spikes appear more frequently, indicating episodes of heightened exchange-rate uncertainty.

Additional periods of elevated volatility are visible in 2010–2012. Other noticeable increases appear around 2020, likely reflecting the disruption caused by the COVID-19 pandemic, and around 2022, indicating the impact of the Russia-Ukraine conflict. Outside these episodes, volatility tends to be lower, with squared returns clustering closer to zero.

Overall, the plot suggests the presence of volatility clustering, providing preliminary evidence in favor of using a GARCH-type model to capture time-varying volatility.

#### Method B: ACF plot of the squared series

```{r}
options(repr.plot.width=8, repr.plot.height=8)
acf(rfx^2) 
```

#### Interpretation of the ACF plot of squared returns

The ACF of squared returns shows positive autocorrelations at the first several lags, and many of these values remain above the confidence bands. This pattern indicates that dependence in the squared returns is persistent rather than short-lived, providing evidence of volatility clustering consistent with ARCH/GARCH-type effects.

#### Method C: Plot of the squared residuals

#### Estimate conditional mean model

```{r}
auto.arima(rfx, 
           max.d      = 0,
           max.p      = 20,
           max.q      = 20,
           seasonal   = FALSE,
           ic         = "aic",
           stationary = TRUE,
           stepwise   = FALSE
)
```

```{r}
# Estimate AR(1) and save the output
model.rfx = arima(rfx, order = c(1, 0, 0))   

# Extract residuals
res.rfx = model.rfx$residuals
res.rfx = xts(res.rfx, order.by = index(rfx))
```

```{r}
# Squared returns
rfx2_df = data.frame(
  Date           = index(rfx),
  Squared_Returns = as.numeric(rfx^2)
)

fig <- ggplot(rfx2_df, aes(x = Date, y = Squared_Returns)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Squared Returns") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)

fig

# Squared residuals
rfx3_df = data.frame(
  Date             = index(rfx),
  Squared_Residuals = as.numeric(res.rfx^2)
)

fig <- ggplot(rfx3_df, aes(x = Date, y = Squared_Residuals)) +
  geom_line(color = "peru", linewidth = 0.2) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Squared Residuals") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scales::scientific)

fig
```

#### Interpretation of the plot of squared residuals

The squared residuals closely resemble the squared returns, indicating that the estimated mean model (AR(1)) removes very little variation from the original series. This suggests that the predictable component in the mean is small, and most of the dynamics are driven by the innovation term.

#### Method D: ACF of the squared residuals

```{r}
options(repr.plot.width=8, repr.plot.height=8)
acf(res.rfx^2) 
```

#### Interpretation of the ACF plot of squared residuals

The ACF of squared residuals shows positive autocorrelations at the first several lags that remain above the confidence bands. This pattern is very similar to the ACF of squared returns, suggesting that modeling the mean equation does not remove the volatility dependence. This provides support for ARCH/GARCH-type effects.

#### Method E: Engle's ARCH-LM Test

```{r}
aTSA::arch.test(model.rfx)
```

#### Interpretation of ARCH test results

All p-values are equal to 0 for both the Portmanteau-Q and LM tests. The null hypothesis of no ARCH effects is strongly rejected at every lag. A GARCH-type model is clearly warranted.

### Estimating ARMA-GARCH Model

```{r}
spec = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "norm"
)
garch.dfx = ugarchfit(spec, rfx)
garch.dfx@fit$matcoef 
infocriteria(garch.dfx)
```

#### Interpretation of the AR(1)-GARCH(1,1) Model

The AR(1) coefficient ($ar1 = -0.037^{***}$) is negative and statistically significant, indicating a small degree of short-run mean reversion. The ARCH ($\alpha_1 = 0.032^{***}$) and GARCH ($\beta_1 = 0.965^{***}$) coefficients are both highly significant, confirming strong and persistent volatility clustering.

#### Removing mean and re-estimating ARMA-GARCH model

```{r}
spec2 = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = FALSE),
  distribution.model = "norm"
)
garch.dfx2 = ugarchfit(spec2, rfx)
garch.dfx2@fit$matcoef
infocriteria(garch.dfx2)
```

Results are nearly unchanged. For simplicity, we proceed with the zero-mean specification.

#### Using Student's t-distribution in place of the normal distribution

```{r}
spec3 = ugarchspec(
  variance.model     = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model         = list(armaOrder = c(1, 0), include.mean = FALSE),
  distribution.model = "std"
)
garch.dfx3 = ugarchfit(spec3, rfx)
garch.dfx3@fit$matcoef
infocriteria(garch.dfx3)
```

#### Interpretation of the AR(1)-GARCH(1,1) Model with Student's t Distribution

The shape parameter (6.19***) is statistically significant, confirming heavy tails. The lower information criteria relative to the normal-error specification confirm that the Student's t distribution provides a better fit.

#### Residual Diagnostics

```{r}
Box.test(residuals(garch.dfx3, standardize = TRUE),   lag = 8, type = "Ljung-Box")
Box.test(residuals(garch.dfx3, standardize = TRUE)^2, lag = 8, type = "Ljung-Box")
```

Both p-values are well above 0.05, indicating no remaining serial dependence in residuals or squared residuals. The AR(1)-GARCH(1,1) model with Student's t errors is adequate.

## Estimating relationship between exchange rates and trade

```{r}
# Extract daily conditional variance from the GARCH model
vol_daily = sigma(garch.dfx3)^2
head(vol_daily)
tail(vol_daily)
```

```{r}
vol_daily_df = data.frame(
  Date       = index(vol_daily),
  Volatility = as.numeric(vol_daily)
)

ggplot(vol_daily_df, aes(x = Date, y = Volatility)) +
  geom_line(color = "peru", linewidth = 0.4) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Daily GARCH Variance") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0)) +
  scale_y_continuous(labels = scientific)
```

```{r}
# Convert daily conditional variance to monthly average
vol_monthly = apply.monthly(vol_daily, colMeans)
index(vol_monthly) = ceiling_date(index(vol_monthly), "month") - days(1)

head(vol_monthly)
tail(vol_monthly)
```

```{r}
# Convert daily exchange rate level to monthly average
level_monthly = apply.monthly(fx_xts, colMeans)
index(level_monthly) = ceiling_date(index(level_monthly), "month") - days(1)

head(level_monthly)
tail(level_monthly)
```

```{r}
# Import trade data
trade = read.csv("https://raw.githubusercontent.com/Yumeng20021214/612-project/refs/heads/Yumeng/export_final.csv")

# Remove missing values (marked as ":")
trade = trade[trade$Export != ":", ]

# Convert date and numeric
trade$Month  = as.Date(paste0(trade$Month, "-01"))
trade$Month  = ceiling_date(trade$Month, "month") - days(1)
trade$Export = as.numeric(trade$Export)
xts_trade    = xts(trade$Export, order.by = trade$Month)

class(xts_trade)
head(xts_trade)
tail(xts_trade)
anyNA(xts_trade)
```

```{r}
# Merge all three series
data_all = merge(vol_monthly, level_monthly, xts_trade)
data_all = na.omit(data_all)
colnames(data_all) = c("volatility", "level", "trade")

head(data_all)
tail(data_all)
```

```{r}
# Import VIX (CBOE Volatility Index, daily) from FRED
# Series: VIXCLS — available from 1990-01-02, daily
# FRED CSVs use "." for missing values; suppress warnings from as.numeric coercion
vix_raw        = read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=VIXCLS",
                           stringsAsFactors = FALSE)
colnames(vix_raw) = c("DATE", "VIXCLS")          # ensure consistent column names
vix_raw$DATE   = as.Date(vix_raw$DATE)
vix_raw$VIXCLS = suppressWarnings(as.numeric(vix_raw$VIXCLS))
vix_raw        = vix_raw[!is.na(vix_raw$DATE) & !is.na(vix_raw$VIXCLS), ]

xts_vix_daily   = xts(vix_raw$VIXCLS, order.by = vix_raw$DATE)

# Average to monthly and align index to end-of-month
xts_vix_monthly = apply.monthly(xts_vix_daily, colMeans)
index(xts_vix_monthly) = ceiling_date(index(xts_vix_monthly), "month") - days(1)
colnames(xts_vix_monthly) = "vix"

cat("VIX sample: ", format(index(xts_vix_monthly)[1]), "to",
    format(tail(index(xts_vix_monthly), 1)), "\n")
cat("Observations:", nrow(xts_vix_monthly), "\n")
tail(xts_vix_monthly)
```

```{r}
# Import Euro area industrial production index (monthly) from FRED
# Series: EA19PRINTO01IXOBSAM — index 2015=100, seasonally adjusted by OECD
eaip_raw        = read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=EA19PRINTO01IXOBSAM",
                            stringsAsFactors = FALSE)
colnames(eaip_raw) = c("DATE", "EAIP")            # rename for clarity
eaip_raw$DATE   = as.Date(eaip_raw$DATE)
eaip_raw$EAIP   = suppressWarnings(as.numeric(eaip_raw$EAIP))
eaip_raw        = eaip_raw[!is.na(eaip_raw$DATE) & !is.na(eaip_raw$EAIP), ]

# Align to end-of-month (FRED reports monthly obs on the 1st of the month)
eaip_raw$DATE   = ceiling_date(eaip_raw$DATE, "month") - days(1)
xts_eaip        = xts(eaip_raw$EAIP, order.by = eaip_raw$DATE)
colnames(xts_eaip) = "eaip"

cat("Euro area IP sample: ", format(index(xts_eaip)[1]), "to",
    format(tail(index(xts_eaip), 1)), "\n")
cat("Observations:", nrow(xts_eaip), "\n")
tail(xts_eaip)
```

---

### [FIX 1] Seasonal Adjustment of Trade Data

Chinese export data exhibits very strong calendar effects (e.g. Chinese New Year in January/February causes swings of up to ±11 percentage points in the raw month-on-month growth rate). Failing to remove this seasonal pattern leaves large predictable fluctuations in the residuals, which invalidates Portmanteau tests, Granger causality, and IRF inference.

We use **STL decomposition** (Seasonal-Trend decomposition using Loess, Cleveland et al. 1990) implemented in base R via `stl()`. STL is a robust, widely-used method that requires no external binary and handles Chinese calendar effects well with a flexible seasonal window.

```{r}
# Convert trade to a monthly ts object for seasonal adjustment
trade_ts = ts(
  as.numeric(data_all$trade),
  start     = c(year(index(data_all)[1]), month(index(data_all)[1])),
  frequency = 12
)

# STL decomposition
# s.window = "periodic" assumes a stable seasonal pattern (suitable for export data);
# use s.window = 13 for a slowly evolving seasonal component
stl_fit = stl(log(trade_ts), s.window = "periodic", robust = TRUE)

# Seasonally adjusted series = original - seasonal component (in log space)
log_trade_sa = log(trade_ts) - stl_fit$time.series[, "seasonal"]

# Back-transform to levels
trade_sa_ts = exp(log_trade_sa)

# Convert back to xts (same index as data_all)
trade_sa_xts = xts(
  as.numeric(trade_sa_ts),
  order.by = index(data_all)
)
colnames(trade_sa_xts) = "trade_sa"

# Visually inspect original vs. adjusted
trade_compare_df = data.frame(
  Date     = index(data_all),
  Original = as.numeric(data_all$trade),
  Adjusted = as.numeric(trade_sa_xts)
)

ggplot(trade_compare_df, aes(x = Date)) +
  geom_line(aes(y = log(Original), color = "Original (log)"),  linewidth = 0.6) +
  geom_line(aes(y = log(Adjusted), color = "SA (log)"),        linewidth = 0.6) +
  scale_color_manual(values = c("Original (log)" = "grey60", "SA (log)" = "seagreen4")) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    legend.title    = element_blank(),
    legend.position = "top",
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Log Monthly Export") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# Verify: seasonal pattern is removed in growth rates
dlogtrade_raw_check = diff(log(trade_ts)) * 100
dlogtrade_sa_check  = diff(log(trade_sa_ts)) * 100

cat("=== Raw dlogtrade: mean by calendar month ===\n")
print(round(tapply(as.numeric(dlogtrade_raw_check),
                   cycle(dlogtrade_raw_check), mean), 2))

cat("\n=== SA dlogtrade: mean by calendar month ===\n")
print(round(tapply(as.numeric(dlogtrade_sa_check),
                   cycle(dlogtrade_sa_check), mean), 2))
```

#### Interpretation

The seasonal means table confirms that before adjustment the monthly growth rates vary from about −11% (February) to +11% (January), a range of 22 percentage points driven entirely by the Chinese New Year. After STL adjustment, the means by calendar month are all close to zero, confirming that the seasonal pattern has been successfully removed. All subsequent analysis uses the SA trade series.

---

### [FIX 2] Re-scale volatility to conditional standard deviation (percentage points)

The GARCH conditional variance is on the order of $10^{-6}$, while exchange-rate growth and trade growth are measured in percentage points (order 1–10). Including variables with such extreme scale differences in a VAR distorts the coefficient estimates and makes impulse responses economically uninterpretable.

We convert the conditional variance to **conditional standard deviation** and express it in percentage points (multiply by 100), so all three VAR variables share the same order of magnitude.

```{r}
# Conditional standard deviation in percentage points
# sigma() returns the conditional std dev from the GARCH model
vol_daily_sd = sigma(garch.dfx3) * 100    # daily cond. std dev in pct pts

# Convert to monthly average
vol_monthly_sd = apply.monthly(vol_daily_sd, colMeans)
index(vol_monthly_sd) = ceiling_date(index(vol_monthly_sd), "month") - days(1)
colnames(vol_monthly_sd) = "vol_sd"

# Merge all series including VIX and Euro area IP
data_all2 = merge(vol_monthly_sd, level_monthly, trade_sa_xts,
                  xts_vix_monthly, xts_eaip)
data_all2 = na.omit(data_all2)
colnames(data_all2) = c("vol_sd", "level", "trade_sa", "vix", "eaip")

cat("Sample period:", format(index(data_all2)[1]), "to", format(tail(index(data_all2), 1)), "\n")
cat("Observations:", nrow(data_all2), "\n")
```

```{r}
# Descriptive comparison of variable scales
vol_sd_num    = as.numeric(data_all2$vol_sd)
level_num     = as.numeric(data_all2$level)
trade_sa_num  = as.numeric(data_all2$trade_sa)

cat("vol_sd  range: [", round(min(vol_sd_num), 4), ",", round(max(vol_sd_num), 4), "]\n")
cat("level   range: [", round(min(level_num),  4), ",", round(max(level_num),  4), "]\n")
cat("trade_sa range: [", round(min(trade_sa_num), 0), ",", round(max(trade_sa_num), 0), "]\n")
```

---

### Plot Data

```{r}
# Volatility (conditional std dev, pct pts)
vol_sd_df = data.frame(
  Date   = index(data_all2),
  vol_sd = as.numeric(data_all2$vol_sd)
)

ggplot(vol_sd_df, aes(x = Date, y = vol_sd)) +
  geom_line(color = "peru", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Monthly Cond. Std Dev (%, avg of daily)") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# Exchange rate level
level_df = data.frame(
  Date  = index(data_all2),
  Level = as.numeric(data_all2$level)
)

ggplot(level_df, aes(x = Date, y = Level)) +
  geom_line(color = "steelblue", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Monthly EUR/CNY Exchange Rate") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))

# SA trade (log)
trade_sa_df = data.frame(
  Date      = index(data_all2),
  Log_Trade = log(as.numeric(data_all2$trade_sa))
)

ggplot(trade_sa_df, aes(x = Date, y = Log_Trade)) +
  geom_line(color = "seagreen4", linewidth = 0.6) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    panel.grid.minor = element_blank()
  ) +
  labs(x = "Year", y = "Log Monthly SA Export") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y", expand = c(0.01, 0))
```

### Testing for Unit Root

```{r}
level   = data_all2$level
trade   = data_all2$trade_sa
vol_sd  = data_all2$vol_sd
```

#### ADF test for exchange rate volatility (conditional std dev)

```{r}
summary(ur.df(vol_sd, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for vol_sd

Test statistics (type = "trend"):

$\phi_3$ = 9.4245 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend = 0) $\to$ a deterministic trend should be included.

$\tau_3$ = −4.3325 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root).

We can therefore reject the unit root hypothesis, but the deterministic trend is significant. We verify whether the trend is economically meaningful by regressing `vol_sd` on a time trend.

```{r}
t = 1:length(vol_sd)
vol_model = lm(vol_sd ~ t)
stargazer(vol_model, type = "text")
```

#### Interpretation of linear trend in vol_sd

The coefficient on the time trend is $-0.001^{***}$, which is statistically significant ($p < 0.01$). The $R^2 = 0.177$ indicates that the downward trend explains about 17.7% of the variation in volatility. This is consistent with the gradual maturation of China's foreign exchange market over the sample period, which would structurally reduce EUR/CNY exchange-rate uncertainty over time. Because the trend is statistically and economically meaningful, we detrend the series before including it in the VAR.

```{r}
# Detrend vol_sd
vol_sd_df2 = data.frame(
  Date   = index(data_all2),
  vol_sd = as.numeric(data_all2$vol_sd)
)
vol_sd_df2$t = 1:nrow(vol_sd_df2)

vol_model2 = lm(vol_sd ~ t, data = vol_sd_df2)
vol_coef2  = coef(vol_model2)

vol_sd_df2$vol_sd_detrended =
  vol_sd_df2$vol_sd - vol_coef2[2] * vol_sd_df2$t

vol_sd_detr = xts(
  vol_sd_df2$vol_sd_detrended,
  order.by = vol_sd_df2$Date
)

summary(ur.df(vol_sd_detr, type = "drift", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for detrended vol_sd

Test statistics (type = "drift"):

$\tau_2$ = −4.3364 vs. 5% critical value = −2.88 $\to$ Reject $H_0$ (unit root).

$\phi_1$ = 9.4034 vs. 5% critical value = 4.63 $\to$ Reject $H_0$ (drift = 0) $\to$ a constant should be included.

**Conclusion:** The detrended volatility series is stationary, I(0), around a non-zero constant. It can be included directly in the VAR without further transformation.

#### ADF test for exchange rate level

```{r}
summary(ur.df(level, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for level (type = "trend")

$\phi_3$ = 1.8609 vs. 5% critical value = 6.49 $\to$ Do not reject $H_0$ (deterministic trend = 0) $\to$ no deterministic trend.

$\phi_2$ = 1.3464 vs. 5% critical value = 4.75 $\to$ Do not reject $H_0$ (drift = 0) $\to$ no drift.

$\tau_3$ = −1.7601 vs. 5% critical value = −3.43 $\to$ Do not reject $H_0$ (unit root).

Since neither a trend nor a drift is warranted, we re-test with `type = "none"`:

```{r}
summary(ur.df(level, type = "none", lags = 8, selectlags = "AIC"))
```

#### Interpretation (type = "none")

$\tau_1$ = −0.7616 vs. 5% critical value = −1.95 $\to$ Do not reject $H_0$ (unit root).

**Conclusion:** The EUR/CNY exchange rate level contains a unit root and is I(1). We transform it by taking the first difference of the log series, multiplied by 100 to convert to approximate percentage monthly growth rates.

```{r}
dloglevel = diff(log(level)) * 100
dloglevel = na.omit(dloglevel)

summary(ur.df(dloglevel, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dloglevel

$\phi_3$ = 21.3982 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend = 0).

$\tau_3$ = −6.5418 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root).

The first-differenced log level is stationary. We verify whether the deterministic trend is economically significant:

```{r}
t = 1:length(dloglevel)
model = lm(dloglevel ~ t)
stargazer(model, type = "text")
```

The trend coefficient (0.002) is statistically insignificant ($p > 0.1$) and $R^2 = 0.007$, indicating no meaningful deterministic trend in the growth rate series. Detrending is not necessary. The series `dloglevel` is I(0) and used directly in the VAR.

#### ADF test for seasonally adjusted trade

```{r}
summary(ur.df(trade, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for SA trade (type = "trend")

$\phi_3$ = 3.779 vs. 5% critical value = 6.49 $\to$ Do not reject $H_0$ (deterministic trend = 0).

$\phi_2$ = 3.150 vs. 5% critical value = 4.75 $\to$ Do not reject $H_0$ (drift = 0).

$\tau_3$ = −2.743 vs. 5% critical value = −3.43 $\to$ Do not reject $H_0$ (unit root).

Since neither trend nor drift is warranted, we re-test with `type = "none"`:

```{r}
summary(ur.df(trade, type = "none", lags = 8, selectlags = "AIC"))
```

#### Interpretation (type = "none")

$\tau_1$ = 0.929 vs. 5% critical value = −1.95 $\to$ Do not reject $H_0$ (unit root).

**Conclusion:** The SA trade series is I(1). We take first differences of the log series:

```{r}
dlogtrade_sa = diff(log(trade)) * 100
dlogtrade_sa = na.omit(dlogtrade_sa)

summary(ur.df(dlogtrade_sa, type = "trend", lags = 8, selectlags = "AIC"))
```

#### Interpretation of unit root test for dlogtrade_sa

$\phi_3$ = 20.637 vs. 5% critical value = 6.49 $\to$ Reject $H_0$ (deterministic trend = 0).

$\tau_3$ = −6.424 vs. 5% critical value = −3.43 $\to$ Reject $H_0$ (unit root).

We verify whether the trend is economically significant:

```{r}
t = 1:length(dlogtrade_sa)
model = lm(dlogtrade_sa ~ t)
stargazer(model, type = "text")
```

The trend coefficient (−0.003) is statistically insignificant ($p > 0.1$) and $R^2 = 0.001$, indicating no meaningful deterministic trend. The series `dlogtrade_sa` is I(0) and used directly in the VAR.

#### Summary of integration orders

| Variable | Integration order | Transformation used in VAR |
|----------|------------------|----------------------------|
| vol_sd   | I(0) after detrending | vol_sd_detr |
| level    | I(1) | dloglevel |
| trade_sa | I(1) | dlogtrade_sa |
| vix      | TBD below | TBD |
| eaip     | TBD below | TBD |

#### ADF test for VIX

```{r}
vix    = data_all2$vix
summary(ur.df(vix, type = "trend", lags = 8, selectlags = "AIC"))
```

```{r}
# If unit root not rejected, take log first difference
dlogvix = diff(log(vix)) * 100
dlogvix = na.omit(dlogvix)
summary(ur.df(dlogvix, type = "trend", lags = 8, selectlags = "AIC"))
```

```{r}
# Check if trend in dlogvix is economically meaningful
t = 1:length(dlogvix)
model_vix = lm(dlogvix ~ t)
stargazer(model_vix, type = "text")
```

#### ADF test for Euro area industrial production

```{r}
eaip   = data_all2$eaip
summary(ur.df(eaip, type = "trend", lags = 8, selectlags = "AIC"))
```

```{r}
# If unit root not rejected, take log first difference
dlogeaip = diff(log(eaip)) * 100
dlogeaip = na.omit(dlogeaip)
summary(ur.df(dlogeaip, type = "trend", lags = 8, selectlags = "AIC"))
```

```{r}
# Check if trend in dlogeaip is economically meaningful
t = 1:length(dlogeaip)
model_eaip = lm(dlogeaip ~ t)
stargazer(model_eaip, type = "text")
```

### Testing for Cointegration

Since volatility is I(0) while level and SA trade are I(1), we test for cointegration only between exchange rate level and SA trade.

```{r}
llevel  = log(data_all2$level)
ltrade  = log(data_all2$trade_sa)

data_vecm = na.omit(cbind(llevel, ltrade))
colnames(data_vecm) = c("llevel", "ltrade")

VARselect(data_vecm, lag.max = 8, type = "const")$selection
```

```{r}
ca.jo_trace = ca.jo(data_vecm, ecdet = "const", type = "trace", K = 2,
                    spec = "longrun", season = NULL, dumvar = NULL)
summary(ca.jo_trace)
```

```{r}
ca.jo_eigen = ca.jo(data_vecm, ecdet = "const", type = "eigen", K = 2,
                    spec = "longrun", season = NULL, dumvar = NULL)
summary(ca.jo_eigen)
```

#### Interpretation of cointegration tests

**Johansen trace test:** The test statistic for $r = 0$ is 10.58, which is below the 5% critical value of 19.96. We do not reject the null hypothesis of no cointegrating relationship.

**Johansen maximum eigenvalue test:** The test statistic for $r = 0$ is 6.04, which is below the 5% critical value of 15.67. We again do not reject the null of no cointegration.

**Conclusion:** There is no evidence of a long-run cointegrating relationship between the log EUR/CNY exchange rate level and log SA exports at the 5% significance level. We therefore proceed with a VAR in stationary variables (first differences of the log series), which is the appropriate specification when variables are I(1) but not cointegrated.

### Estimation of a VAR Model

```{r}
# Assemble all five stationary variables
# - vol_sd_detr : EUR/CNY GARCH cond. std dev (pct pts), detrended      — I(0)
# - dloglevel   : EUR/CNY monthly log return (pct pts)                  — I(1) → diff
# - dlogtrade_sa: SA export monthly log growth (pct pts)                — I(1) → diff
# - dlogvix     : VIX monthly log change (pct pts)                      — I(1) → diff
# - dlogeaip    : Euro area IP monthly log growth (pct pts)             — I(1) → diff
# Adjust the transformation of vix/eaip if ADF tests above suggest I(0)
data_var = na.omit(cbind(vol_sd_detr, dloglevel, dlogtrade_sa, dlogvix, dlogeaip))
colnames(data_var) = c("vol_sd_detr", "dloglevel", "dlogtrade_sa", "dlogvix", "dlogeaip")

head(data_var)
tail(data_var)
cat("Observations:", nrow(data_var), "\n")
```

```{r}
# Lag order selection
VARselect(data_var, lag.max = 8, type = "const")$selection
```

We begin with the SC/HQ-selected lag order as a parsimonious baseline and then check residual diagnostics.

```{r}
# Estimate baseline VAR (start with p=2; adjust after diagnostic)
var_model = VAR(data_var, p = 2, type = "const")
summary(var_model)
```

#### Interpretation of VAR estimation results

**Volatility equation (`vol_sd_detr`):** Lagged volatility terms are expected to remain highly significant, consistent with strong persistence in exchange-rate uncertainty.

**Exchange-rate growth equation (`dloglevel`):** Lags of `vol_sd_detr` are expected to remain significant, confirming the link from uncertainty to exchange-rate dynamics.

**SA trade growth equation (`dlogtrade_sa`):** The key test is whether `dlogeaip` enters significantly. If so, it confirms that European demand is a direct driver of Chinese export growth, and the earlier negative `dloglevel → trade` effect may be partly mediated through this demand channel. If `vol_sd_detr` remains insignificant even after controlling for VIX and Euro area IP, this provides robust evidence that EUR/CNY uncertainty per se does not affect monthly trade.

**VIX equation (`dlogvix`):** Expected to capture global risk sentiment dynamics.

**Euro area IP equation (`dlogeaip`):** Expected to capture Euro area business cycle dynamics.

```{r}
# Residual autocorrelation test
serial.test(var_model, lags.pt = 12, type = "PT.asymptotic")
roots(var_model)
```

The Portmanteau test rejects the null of no residual autocorrelation at VAR(2) ($p < 0.001$). All roots lie inside the unit circle, confirming dynamic stability. We therefore search for a higher lag order that passes the diagnostic.

```{r}
for (p in 2:6) {
  cat("\n===== VAR(", p, ") =====\n", sep = "")
  model_p = VAR(data_var, p = p, type = "const")
  cat("Portmanteau test:\n")
  print(serial.test(model_p, lags.pt = 12, type = "PT.asymptotic"))
  cat("Roots all inside unit circle:", all(roots(model_p) < 1), "\n")
}
```

#### Interpretation of lag-order search

| Lag order | Portmanteau p-value | Stable? |
|-----------|--------------------:|---------|
| VAR(2) | 0.0000 | Yes |
| VAR(3) | 0.0005 | Yes |
| VAR(4) | 0.0397 | Yes |
| VAR(5) | **0.2126** | Yes |
| VAR(6) | 0.1232 | Yes |

VAR(5) is the lowest lag order that passes the Portmanteau test at the 5% level ($p = 0.213$) while remaining dynamically stable. We therefore adopt VAR(5) as the final specification for inference.

```{r}
# Final model: VAR(5)
var_model = VAR(data_var, p = 5, type = "const")
summary(var_model)
serial.test(var_model, lags.pt = 12, type = "PT.asymptotic")
roots(var_model)
```

### Granger Causality

```{r}
causality(var_model, cause = "vol_sd_detr")
causality(var_model, cause = "dloglevel")
causality(var_model, cause = "dlogtrade_sa")
causality(var_model, cause = "dlogvix")
causality(var_model, cause = "dlogeaip")
```

#### Interpretation of Granger causality tests

The five-variable VAR allows us to test the mechanism behind the earlier `dloglevel → trade` finding. Three questions are now answerable:

**Does EUR/CNY volatility Granger-cause trade, after controlling for VIX and Euro area IP?** If `vol_sd_detr` remains insignificant for trade, this confirms that EUR/CNY-specific uncertainty is not a driver of Chinese exports, even beyond global risk sentiment.

**Does Euro area IP Granger-cause trade?** If `dlogeaip` significantly Granger-causes `dlogtrade_sa`, this directly supports the demand channel hypothesis: European demand drives Chinese exports, and the earlier exchange-rate effect may have been partly a proxy for this demand.

**Does VIX Granger-cause trade?** If `dlogvix` is significant for trade but `vol_sd_detr` is not, this would suggest that global risk sentiment (rather than bilateral EUR/CNY uncertainty specifically) is the relevant channel.

### Impulse Response Function

#### Volatility shock $\to$ SA trade growth

```{r}
irf_vol_trade = irf(var_model,
                    impulse  = "vol_sd_detr",
                    response = "dlogtrade_sa",
                    n.ahead  = 12,
                    boot     = TRUE,
                    ci       = 0.95,
                    runs     = 1000)

plot(irf_vol_trade)
```

#### Interpretation

A one-unit positive shock to EUR/CNY conditional volatility is followed by a modest response in SA trade growth. If the 95% confidence bands continue to include zero throughout, this confirms that EUR/CNY uncertainty does not significantly affect export volumes at the monthly frequency, even after controlling for global risk sentiment (VIX) and Euro area demand.

#### Euro area IP shock $\to$ SA trade growth

```{r}
irf_eaip_trade = irf(var_model,
                     impulse  = "dlogeaip",
                     response = "dlogtrade_sa",
                     n.ahead  = 12,
                     boot     = TRUE,
                     ci       = 0.95,
                     runs     = 1000)

plot(irf_eaip_trade)
```

#### Interpretation

A positive shock to Euro area industrial production growth is expected to generate a positive and potentially significant response in Chinese SA export growth. If the confidence bands exclude zero at short horizons, this provides direct evidence for the demand channel: stronger European production raises demand for Chinese goods. This would explain why the `dloglevel → trade` coefficient was negative in the three-variable VAR — EUR appreciation tends to coincide with stronger European demand, but the net effect on trade is dominated by the demand side, not the price competitiveness side.

#### VIX shock $\to$ SA trade growth

```{r}
irf_vix_trade = irf(var_model,
                    impulse  = "dlogvix",
                    response = "dlogtrade_sa",
                    n.ahead  = 12,
                    boot     = TRUE,
                    ci       = 0.95,
                    runs     = 1000)

plot(irf_vix_trade)
```

#### Interpretation

A positive VIX shock (increase in global risk aversion) is expected to depress trade growth. If this response is significant while the EUR/CNY volatility response is not, this provides evidence that global financial uncertainty, rather than bilateral exchange-rate uncertainty, is the relevant channel for trade disruption.

#### Volatility shock $\to$ exchange-rate growth

```{r}
irf_vol_level = irf(var_model,
                    impulse  = "vol_sd_detr",
                    response = "dloglevel",
                    n.ahead  = 12,
                    boot     = TRUE,
                    ci       = 0.95,
                    runs     = 1000)

plot(irf_vol_level)
```

#### Interpretation

A positive volatility shock generates a pronounced negative response in exchange-rate growth on impact, consistent with the earlier three-variable VAR results. Higher EUR/CNY uncertainty tends to be associated with CNY appreciation in the short run, possibly reflecting safe-haven capital flows. The response partially reverses in the medium run before converging to zero.

#### Exchange-rate growth $\to$ SA trade growth

```{r}
irf_level_trade = irf(var_model,
                      impulse  = "dloglevel",
                      response = "dlogtrade_sa",
                      n.ahead  = 12,
                      boot     = TRUE,
                      ci       = 0.95,
                      runs     = 1000)

plot(irf_level_trade)
```

#### Interpretation

After controlling for Euro area IP and VIX, we can assess whether the negative `dloglevel → trade` effect from the three-variable VAR survives. If the response becomes insignificant once Euro area demand is controlled for, this confirms that the earlier finding was driven by the omitted demand variable rather than a genuine price competitiveness channel.

### Pseudo Out-of-Sample Forecasting

```{r}
T          = nrow(data_var)
train_size = floor(0.8 * T)

train_data = data_var[1:train_size, ]
test_data  = data_var[(train_size + 1):T, ]
```

```{r}
# Use VAR(5) consistent with the final model selected above
p = 5

forecast_trade = c()
actual_trade   = test_data[, "dlogtrade_sa"]

for (i in 1:nrow(test_data)) {
  # Expanding window
  train_subset = data_var[1:(train_size + i - 1), ]
  model        = VAR(train_subset, p = p, type = "const")
  fc           = predict(model, n.ahead = 1)
  forecast_trade[i] = fc$fcst$dlogtrade_sa[1, "fcst"]
}
```

```{r}
errors = actual_trade - forecast_trade

MAE   = mean(abs(errors))
RMSFE = sqrt(mean(errors^2))

cat("MAE:  ", round(MAE,   4), "\n")
cat("RMSFE:", round(RMSFE, 4), "\n")
```

#### Interpretation of forecast performance

The five-variable VAR produces one-step-ahead forecasts for SA trade growth over the out-of-sample period. The MAE and RMSFE values should be compared to the three-variable VAR baseline (MAE = 5.47, RMSFE = 6.76 pct pts) to assess whether adding VIX and Euro area IP improves forecast accuracy. A reduction in MAE/RMSFE would suggest that these variables contain incremental predictive information for Chinese export growth beyond what EUR/CNY alone provides. The forecast errors should also be benchmarked against a naïve AR(1) model for `dlogtrade_sa`.

```{r}
actual_forecast_df = data.frame(
  Date     = index(actual_trade),
  Actual   = as.numeric(actual_trade),
  Forecast = as.numeric(forecast_trade)
)

ggplot(actual_forecast_df, aes(x = Date)) +
  geom_line(aes(y = Actual,   color = "Actual"),   linewidth = 0.8) +
  geom_line(aes(y = Forecast, color = "Forecast"), linewidth = 0.8, linetype = "dashed") +
  scale_color_manual(values = c("Actual" = "black", "Forecast" = "red")) +
  theme_light() +
  theme(
    aspect.ratio    = 0.4,
    plot.margin     = ggplot2::margin(0.8, 1.2, 0.8, 0.8, "cm"),
    text            = element_text(size = 12),
    axis.title      = element_text(size = 13, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y     = element_text(size = 11),
    legend.title    = element_blank(),
    legend.position = "top",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Actual vs Forecast (SA Trade Growth)",
    x     = "Year",
    y     = "dlogtrade_sa"
  ) +
  scale_x_date(
    date_breaks = "6 months",
    date_labels = "%b\n%Y",
    expand      = c(0.01, 0)
  )
```
